Data preparation report

### 2023-06-09 00_dataPrepReport.Rmd
changed data cleaning threshold to 30 days instead of 1/3 of time

```{r remove-cages}
# Remove cages using Marta's data
# Identify and remove capture dates using Marta's code --------------------
## To identify the capture dates, first we need to classify the roosts (now using the get_roosts() function). Then, if the bird roosted within 500 m of the capture site, it was considered to be captured and that day, and the following day, are excluded from the dataset. This has been validated. Note: this protocol doesn't work for the Carmel because of the position of the roost sites relative to cages. But we're just dealing with southern individuals, so that's okay.
#roosts <- get_roosts_df(df = datAnnot, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
#save(roosts, file = "data/roosts.Rda")
load("data/roosts.Rda")

@@

for(i in 1:nrow(crds)){ # for each roost point...
  DistanceMat[i,] <- round(geosphere::distm(crds[i,], sites[,c(3,2)]), 2) # calculate distance using geosphere::distm
  print(i)
}

@@

The resulting data has dropped from **`r nrow(datAnnot_noCaptures)`** to
**`r nrow(datAnnot)`** rows, a
**`r round((nrow(datAnnot_noCaptures)-nrow(datAnnot))*100/nrow(datAnnot_noCaptures), 3)`% decrease**.
It has `r ncol(datAnnot)` columns and
**`r length(unique(datAnnot$Nili_id))`** unique individuals.

@@

datAnnot2 <- datAnnot_noCaptures %>%
  dplyr::select(-c("sex")) %>%
  left_join(as, by = "Nili_id")
nrow(datAnnot2) == nrow(datAnnot)
length(unique(datAnnot2$Nili_id)) # 125
```

@@

```{r clean-data}
# Clean the data
## Region masking, downsampling, removal of speed outliers, setting altitude outliers to NA, etc.
mask <- sf::st_read("data/CutOffRegion.kml")
# datAnnotCleaned <- vultureUtils::cleanData(dataset = datAnnot, mask = mask, inMaskThreshold = 0.33, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T) %>% st_drop_geometry()
# save(datAnnotCleaned, file = "data/datAnnotCleaned.Rda")
load("data/datAnnotCleaned.Rda")
dim(datAnnotCleaned) # about 27 percent of data outside the israel region (roughly)
length(unique(datAnnotCleaned$Nili_id))
```

@@

```{r get-roosts}
# Get roosts for each season ----------------------------------------------
# forRoosts <- map(seasons, ~.x %>% sf::st_as_sf(coords = c("location_long", "location_lat"), remove = FALSE))
# roosts_seasons <- purrr::map(forRoosts, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id")) 
# roosts_seasons <- roosts_seasons %>%
#   map(., ~st_as_sf(.x, crs = "WGS84", coords = c("location_long", "location_lat")))
# save(roosts_seasons, file = "data/roosts_seasons.Rda")
load("data/roosts_seasons.Rda")
```

>>>>

```{r remove-cages}
# Remove cages using Marta's data
# Identify and remove capture dates using Marta's code --------------------
## To identify the capture dates, first we need to classify the roosts (now using the get_roosts() function). Then, if the bird roosted within 500 m of the capture site, it was considered to be captured and that day, and the following day, are excluded from the dataset. This has been validated. Note: this protocol doesn't work for the Carmel because of the position of the roost sites relative to cages. But we're just dealing with southern individuals, so that's okay.
roosts <- get_roosts_df(df = datAnnot, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
save(roosts, file = "data/roosts.Rda")
load("data/roosts.Rda")

@@

for(i in 1:nrow(crds)){ # for each roost point...
  DistanceMat[i,] <- round(geosphere::distm(crds[i,], sites[,c(3,2)]), 2) # calculate distance using geosphere::distm
  #print(i)
}

@@

The resulting data has dropped from **`r nrow(dat2)`** to
**`r nrow(datAnnot_noCaptures)`** rows, a
**`r round((nrow(dat2)-nrow(datAnnot_noCaptures))*100/nrow(dat2), 3)`% decrease**.
It has `r ncol(dat2)` columns and
**`r length(unique(datAnnot_noCaptures$Nili_id))`** unique individuals.

@@

datAnnot2 <- datAnnot_noCaptures %>%
  dplyr::select(-c("sex")) %>%
  left_join(as, by = "Nili_id")
nrow(datAnnot2) == nrow(datAnnot_noCaptures)
length(unique(datAnnot2$Nili_id)) # 125
```

@@

```{r clean-data, echo = T, include = T}
# Clean the data
## Region masking, downsampling, removal of speed outliers, setting altitude outliers to NA, etc.
mask <- sf::st_read("data/CutOffRegion.kml")
datAnnotCleaned <- vultureUtils::cleanData(dataset = datAnnot, mask = mask, inMaskThreshold = 30, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T) %>% st_drop_geometry()
save(datAnnotCleaned, file = "data/datAnnotCleaned.Rda")
load("data/datAnnotCleaned.Rda")
```

@@

```{r get-roosts}
# Get roosts for each season ----------------------------------------------
# forRoosts <- map(seasons, ~.x %>% sf::st_as_sf(coords = c("location_long", "location_lat"), remove = FALSE))
# roosts_seasons <- purrr::map(forRoosts, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id")) 
# roosts_seasons <- roosts_seasons %>%
#   map(., ~st_as_sf(.x, crs = "WGS84", coords = c("location_long", "location_lat"), remove = F))
# save(roosts_seasons, file = "data/roosts_seasons.Rda")
load("data/roosts_seasons.Rda")
```
